%%% Введение
\begin{Russian}
\section{Введение}
\label{sec:intro}

Данное исследование посвящено проведению натурного эксперимента по моделированию заражения компьютерным вирусом в эмуляторах Kathara и Containerlab. Изучение механизмов распространения вируса позволяет проанализировать возможности систем, в которых проводятся исследования, и их устойчивость к вредоносному трафику. Традиционно для исследования динамики распространения вирусов применяются математические модели, заимствованные из эпидемиологии ~\cite{Hoang2025}, такие как модели SIR (Susceptible-Infected-Recovered), SEIR (Susceptible-Exposed-Infected-Recovered) и SEIRS (Susceptible-Exposed-Infected-Recovered-Susceptible). Эти модели позволяют описывать процессы заражения, инкубации и восстановления в популяции узлов сети.

Средство моделирования Kathara дает возможность создавать виртуальные сетевые топологии и исследовать их поведение в контролируемых условиях. Containerlab широко применяется в индустрии для тестирования и валидации сетевых решений ~\cite{Huu2025} ~\cite{Usmanov2018}. Обе платформы основываются на Docker контейнерах и способны эмулировать сложные сетевые топологии, однако прямая сравнительная оценка их возможностей в контексте моделирования распространения компьютерного вируса до сих пор не была проведена. Эта работа направлена на оценку возможностей систем при работе с различными эпидемиологическими моделями, а также на понимание ограничений платформ с точки зрения точности результатов.

Использование хорошо изученных эпидемиологических моделей SIR, SEIR и SEIRS в качестве теоретической основы позволяет сосредоточить внимание именно на системных параметрах эмуляторов: длительность эмуляции, потреблении памяти и метриках сетевого трафика. Исследование направлено на изучение поведения средств натурного моделирования в решении задач кибербезопасности.

\end{Russian}

\begin{English}
\section{Introduction}
\label{sec:intro}

This study is devoted to conducting a field experiment on modeling computer virus infection in Kathara and Containerlab emulators. Studying the mechanisms of virus propagation allows us to analyze the capabilities of the systems in which the research is conducted and their resistance to malicious traffic. Traditionally, mathematical models borrowed from epidemiology ~\cite{Hoang2025}, such as SIR (Susceptible-Infected-Recovered) and SEIR (Susceptible-Exposed-Infected-Recovered), and SEIRS (Susceptible-Exposed-Infected-Recovered-Susceptible) models. These models describe the processes of infection, incubation, and recovery in a population of network nodes.

The Kathara simulation tool enables the creation of virtual network topologies and the investigation of their behavior under controlled conditions. Containerlab is widely used in the industry for testing and validating network solutions ~\cite{Huu2025} ~\cite{Usmanov2018}. Both platforms are based on Docker containers and are capable of emulating complex network topologies, but a direct comparative assessment of their capabilities in the context of modeling the spread of a computer virus has not yet been conducted. This work aims to evaluate the capabilities of the systems when working with different epidemiological models, as well as to understand the limitations of the platforms in terms of the accuracy of the results.

The use of well-studied epidemiological models SIR, SEIR, and SEIRS as a theoretical basis allows us to focus specifically on the system parameters of emulators: emulation duration, memory consumption, and network traffic metrics. The study aims to investigate the behavior of full-scale simulation tools in solving cybersecurity problems.

\end{English}

%%% Структура статьи
%\begin{Russian}
%\subsection{Структура статьи}

%В разделе~\ref{sec:models} вводятся уравнения моделей распространения компьютерного вируса. В разделе~\ref{sec:modelling} приводится краткое описание топологий и программной реализации эксперимента.

%\end{Russian}

%\begin{English}
%\subsection{Structure of the paper}
%\end{English}
%\label{sec:structure}

% \begin{English}
% In paragraph~\ref{sec:notation} we prosecuted provides basic notation and conventions
% used in the article. In paragraph~\ref{sec:maxwell_curv} are the main
% relations for the Maxwell's equations in curvilinear coordinates (for
% more detailed discussion the reader can be refer to other 
% authors articles~\cite{kulyabov:2012:vestnik:2012-1}). In paragraph~\ref{sec:formal_geometr} are presented actual
% calculations on Plebanski geometrization.
% \end{English}

%%% Обозначения и соглашения
%\begin{Russian}
%\subsection{Обозначения и соглашения}

%\begin{enumerate}

%\item Будем придерживаться следующих соглашений.  Греческие индексы ($\beta$, $\sigma$, $\gamma$, $\xi$) будут описывать вероятности заражения, инкубационного периода, выздоровления и потери иммунитета соответственно.

%\end{enumerate}
%\end{Russian}

%\begin{English}
%\subsection{Notations and conventions}
%\end{English}
%\label{sec:notation}




% \begin{English}
%   \begin{enumerate}

%   \item We will use the notation of abstract
%     indices~\cite{penrose-rindler:spinors::en}. In this notation tensor
%     as a
%     complete object is denoted merely by an index (e.g., $x^{i}$). Its
%     components are
%     designated by underlined indices (e.g., $x^{\crd{i}}$).
    
%   \item We will adhere to the following agreements. Greek indices
%     ($\alpha$, $\beta$) will refer to the four-dimensional space, in the
%     component form it looks like: $\crd{\alpha} = \overline{0,3}$.  Latin
%     indices from the middle of the alphabet ($i$, $j$, $k$) will refer
%     to the three-dimensional space, in the component form it looks like:
%     $\crd {i} = \overline{1,3}$.
    
%   \item The comma in the index denotes a partial derivative with respect to
%     corresponding coordinate ($f_{, i} := \partial_{i}f$); semicolon
%     denotes a covariant derivative ($f_{;i} := \nabla_{i} f$).

%   \item The CGS symmetrical system~\cite{sivukhin:1979:ufn::en} is used for 
%   notation of the equations of electrodynamics.

%   \end{enumerate}
% \end{English}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End:
